\vspace{-2cm}A generative model for urban firm networks

### 2022-05-16 main.tex
Update on Overleaf.

\begin{flushleft}
{\Large
\textbf\newline{Modelling growth of urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juste Raimbault\textsuperscript{1,2,3 *},
Natalia Zdanowska\textsuperscript{3,4,5},
Elsa Arcaute\textsuperscript{4}
\\
\bigskip
\textbf{1} LASTIG, Univ Gustave Eiffel, IGN-ENSG, Saint-Mand{\'e}, France
\\
\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
\\
\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
\\
\textbf{4} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
\\
\textbf{5} Luxembourg Institute of Socio-Economic Research, Esch-sur-Alzette, Luxembourg

@@

%Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}. At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}. At country level, trade networks have been extensively studied from the complex networks perspective \cite{garlaschelli2005structure}, but these are generally not explored at a more granular level due to the lack of city-level data. Within regional science, input-output models are used as generative urban models of flows between geographical areas \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect} at different geographical scales.

Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}. At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}. At country level, trade networks have been extensively studied from the complex networks perspective \cite{garlaschelli2005structure}, but these are generally not explored at a more granular level due to the lack of city-level data. 
Within this work, we contribute to the research at an interurban level, exploring the position and dynamics of cities in the socio-economic world-wide space \cite{pumain2018evolutionary} \cite{berry1964cities}, by considering the interactions between the firms in the cities. 
An extensive literature exists exploring firm-to-firm level networks in economic geography. For example, within the input-output models framework, these are used to study inter-sectors and inter-region transactions \cite{Rodriguesetal2016}, or firm level productivity, to understand macroeconomic level outcomes \cite{Grassi2016} and economic resilience \cite{Ialongo2021}. %Studies of these firm networks at city level remain still scarce. 
At a more granular level, urban firm networks have been investigated in complexity science \cite{RozenblatNeal2021} to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}. Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}. This work complements this effort by proposing a generative mechanism of such networks.

@@

The specific question we tackle is the complementarity of the different processes to generate urban firm networks. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinants of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}. More precisely : (i) we firstly introduce a generative network model to simulate the growth of such linkages at urban areas scale, combining multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the source and target nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distances; (ii) the model allows us to then compare the effect of the different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; (iii) we develop an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018};  (iv) we use this data to calibrate the model on the empirical network; and (v) we explore an application of the model to an economic shock scenario, with pandemic-like stylised facts.

%The specific question we tackle is the complementarity of different processes to generate urban firm networks. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinants of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}. More precisely : (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the origin and destination nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.

@@

\subsubsection*{Rationale}
%The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. 
The model is expected to capture a single macroscopic urban scale, which corresponds to the level at which nodes are cities. These are characterised by their economic size and their industrial sector profile. %The links in the network are constructed from the relationship between these characteristics and hence induced from their underlying firms. Different distances, or proximity measures, play a role in establishing linkages, as pointed out by \cite{martinus2018global}.
Links are then created considering these features and different distance, or proximity measures \cite{martinus2018global}. 
Note that, geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. 
We therefore combine several features driving network growth: (i) geographical proximity (in our case the crow-fly distance between cities, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (this is confirmed to be  relevant by the fixed effect regression described below); (iii) city economic size, which is known to give rise to agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages, which capture the influence of existing linkages created in the past. This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.

@@

When focusing on the community structure of the network, different modularity maximisation algorithms (ran on the undirected corresponding network) result in the same undirected modularity of 0.38, a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and of 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp. 15). In comparison, when taking as communities 29 existing countries, the (directed?)  modularity is 0.32. This indicates that the structure of flows present certain regional patterns. Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities. We show in Fig.~\ref{fig:fig2} a map of the communities obtained.

>>>>

\begin{flushleft}
{\Large
\textbf\newline{Modelling growth of urban firm networks}
}
\newline

\\
Juste Raimbault\textsuperscript{1,2,3 *},
Natalia Zdanowska\textsuperscript{3,4,5},
Elsa Arcaute\textsuperscript{4}
\\
\bigskip
\textbf{1} LASTIG, Univ Gustave Eiffel, IGN-ENSG, Saint-Mand{\'e}, France
\\
\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
\\
\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
\\
\textbf{4} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
\\
\textbf{5} Luxembourg Institute of Socio-Economic Research, Esch-sur-Alzette, Luxembourg

@@

Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}. At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}. At country level, trade networks have been extensively studied from the complex networks perspective \cite{garlaschelli2005structure}, but these are generally not explored at a more granular level due to the lack of city-level data. 
Within this work, we contribute to the research at an interurban level, exploring the position and dynamics of cities in the socio-economic world-wide space \cite{pumain2018evolutionary} \cite{berry1964cities}, by considering the interactions between the firms in the cities. 
An extensive literature exists exploring firm-to-firm level networks in economic geography. For example, within the input-output models framework, these are used to study inter-sectors and inter-region transactions \cite{Rodriguesetal2016}, or firm level productivity, to understand macroeconomic level outcomes \cite{Grassi2016} and economic resilience \cite{Ialongo2021}.
At a more granular level, urban firm networks have been investigated in complexity science \cite{RozenblatNeal2021} to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}. Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}. This work complements this effort by proposing a generative mechanism of such networks.

@@

The specific question we tackle is the complementarity of the different processes to generate urban firm networks. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinants of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}. More precisely: (i) we firstly introduce a generative network model to simulate the growth of such linkages at urban areas scale, combining multiple factors influencing link formation, such as the economic intensity of the urban areas corresponding to the source and target nodes, the industrial sector proximity, the strength of previous links, and the geographical and socio-cultural distances; (ii) the model allows us to then compare the effect of the different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; (iii) we develop an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018};  (iv) we use this data to calibrate the model on the empirical network; and (v) we explore an application of the model to an economic shock scenario, with pandemic-like stylised facts.

@@

\subsubsection*{Rationale}

The model is expected to capture a single macroscopic urban scale, which corresponds to the level at which nodes are cities. These are characterised by their economic size and their industrial sector profile. Links are then created considering these features and different distance, or proximity measures \cite{martinus2018global}. Note that, geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. We therefore combine several features driving network growth: (i) geographical proximity (in our case the crow-fly distance between cities, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (this is confirmed to be  relevant by the fixed effect regression described below); (iii) city economic size, which is known to give rise to agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages, which capture the influence of existing linkages created in the past. This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.

@@

To focus on the community structure of the network, we test different modularity maximisation algorithms ran on the undirected corresponding network. The Louvain algorithm \cite{blondel2008fast} provides an undirected modularity of $0.388$, with 19 communities. The corresponding directed modularity computed following \cite{nicosia2009extending} is $0.362$. The greedy algorithm of \cite{clauset2004finding} results in 35 communities for a undirected modularity of $0.376$ and an directed modularity of $0.352$. In comparison, when taking as communities 29 existing countries, the corresponding directed modularity is $0.323$. This indicates that the structure of flows present certain regional patterns. Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.050 \pm 0.0017$, confirming the statistical significance of these communities. We show in Fig.~\ref{fig:fig2} a map of the communities obtained.
